Flare Rates and Rotation Periods of 5700 Stars Younger than 300 Myr

### 2023-11-02 src/tex/ms.tex
power law fits to Ro and flare rates

\section{Flare Rates Compared to P$_{rot}$}

\begin{figure}[ht!]
    \script{prot_histograms.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/prot_histograms.pdf}
        \caption{
            Rotation periods and flare rates for M, K, and G stars (top to bottom).
        }
        \label{fig:prot_histograms}
    \end{centering}
\end{figure}

@@

We measured the flare rates of planet hosting stars $< 300$\,Myr, comparable to the ages of our primary sample.
We followed the methods outlined in Section~\ref{sec:methods} to detect and vet flares. For each star, we created
an equivalent sample with respect to both age and $T_\textrm{eff}$ to compare the flare rates too. We considered
stars with ages $\pm 30$\,Myr of the planet hosting star and $T_\textrm{eff} \pm 1000$\,K. We calculate the overall
flare rate as described in \cite{Feinstein22}:

\begin{equation}
  \mathcal{R} = \frac{1}{\tau_\textrm{obs}} \left( \sum_{i=1}^{N} p_i \right)
\end{equation}

where $\mathcal{R}$ is the flare rate in day$^{-1}$, $\tau_\textrm{obs}$ is the total amount of time a target was
observed with TESS, and $p_i$ is the probability that flare $i$ is a true flare as assigned by \texttt{stella}. We present
the flare rates of planet-hosting stars and a comparable sample of stars in Figure~\ref{fig:yp_rates} and report the measured
rates in Table~\ref{tab:yp_rates}. For the comparable sample, we report the median flare rate, and the lower 16\textsuperscript{th}
and upper 84\textsuperscript{th} percentiles.

>>>>

\section{Flare Rates Compared to $R_0$}

The Rossby number, $R_0$, is a term which marginalizes over several properties which are known to
affect the stellar dynamo, such as the rotation period and stellar mass. We convert our measured
rotation periods to $R_0$, which is defined as $R_0 = P_\textrm{rot}/\tau$, where $\tau$ is the
convective turnover time. We approximate $\tau$ following the prescription in \cite{wright11}. We
equate the flare rates for individual stars as

\begin{equation}\label{eqn:fr}
  \mathcal{R} = \frac{1}{t_\textrm{obs}} \left( \sum_{i=1}^{N} p_i \right)
\end{equation}

where $\mathcal{R}$ is the flare rate in units of day$^{-1}$, $t_\textrm{obs}$ is the total amount of time a target was
observed with TESS, and $p_i$ is the probability that flare $i$ is a true flare as assigned by \texttt{stella}.
We compare the calculated $R_0$ to measured flare rates for all stars we measured $P_\textrm{rot}$ for. The
results are presented in Figure~\ref{fig:prot_histograms}.

\begin{figure*}[ht!]
    \script{prot_histograms.py}
    \begin{centering}
        \includegraphics[width=\textwidth]{figures/prot_histograms.pdf}
        \caption{
            Comparison of Rossby Number, $R_0$ and flare rate for young GKM stars.
            We find no change in the average flare rate for GKM stars $< 250$\,Myr.
            We see evidence that as $R_0$ increases, the average flare rate decreases.
            This is similar to results presented in \cite{medina20}, however our sample extends this
            trend to young GKM stars, while previous results were limited to nearby M dwarfs. The top row are
            stars with ages $4.5 - 50$\,Myr; the bottom row are stars with ages $50 - 250$\,Myr. The
            histograms are colored by number of stars in each bin.
        }
        \label{fig:prot_histograms}
    \end{centering}
\end{figure*}

We evaluate the relationship between age, spectral type, flare rate, and $R_0$. We divide the sample between stars
younger and older than 50~Myr, which roughly correlates to the age at which GKM stars turn onto the main sequence.
For stars younger than 50~Myr, we find the flare rate slightly decreases with increasing $R_0$, although there is a
significant amount of scatter in this relationship. For stars between $50 - 250$~Myr, we see this relationship much
clearer. For M stars, we see minimal evolution in both average flare rate and $R_0$ between the two samples. For K
stars, we see $R_0$ evolving over the first 250~Myr, while the scatter in the flare rate decreases. For G stars, we
see the scatter in $R_0$ decreasing, and the average flare rate across the sample decreases. We present a compiled histogram
for all stars in our sample in the right column of Figure~\ref{fig:prot_histograms}.

To better understand this trend, we fit three types of functions: (i) a constant value, (ii) a single power law, and
(iii) a piece-wise function consisting of a constant value and a power law. For (iii), we computed these fits across a range
of $R_0$ for where the turnover should occur. We binned the cumulative dataset consisting of all GKM stars into 50 evenly-spaced bins
in log-space from $log_{10}(R_0) = [-2, 0]$. We computed the $\chi^2$ between
each of these fits and the data. For stars $4.5 - 50$~Myr, we find the distribution is best-fit with a single
power law with slope $m = -0.243 \pm 0.043$ and y-intercept $b = -1.221 \pm 0.061$. Converting the $\chi^2$ to a standard
deviation, we find a single fit is preferred by $\sim 3\sigma$. For stars $50 - 250$~Myr, we find the distribution
is best-fit with a piece-wise function of the form

\begin{equation}
  \mathcal{R} =
  \Bigg \{
  \begin{array}{ll}
        C & R_0 \leq 0.15 \\
        10^b * R_0^m  & R_0 > 0.15 \\
  \end{array}
\end{equation}

where $\mathcal{R}$ is the flare rate, $C = 0.126 \pm 0.006$, $m = -0.986 \pm 0.119$, and $b = -1.687 \pm 0.078$.
We test how our bin size affected our $R_0$ turnover. To do this, we refit our data assuming binning from $20 - 100$
bins and find $R_0$ to be consistent with $R_0 = 0.15 \pm 0.02$. The location of the turnover is consistent to
within $1\sigma$ with what has been seen in other observations of magnetic saturation for partially and
fully convective stars \citep[e.g. $L_X/L_\textrm{bol}$; ][]{wright18}.

@@

We measured the flare rates of planet hosting stars $< 300$\,Myr, comparable to the ages of our primary sample.
We followed the methods outlined in Section~\ref{sec:methods} to detect and vet flares. For each star, we created
an equivalent sample with respect to both age and $T_\textrm{eff}$ to compare the flare rates too. We considered
stars with ages $\pm 30$\,Myr of the planet hosting star and $T_\textrm{eff} \pm 1000$\,K. We calculate the flare rate
following Equation~\ref{eq:fr}. We present
the flare rates of planet-hosting stars and a comparable sample of stars in Figure~\ref{fig:yp_rates} and report the measured
rates in Table~\ref{tab:yp_rates}. For the comparable sample, we report the median flare rate, and the lower 16\textsuperscript{th}
and upper 84\textsuperscript{th} percentiles.
